\singlespacing \vspace{-1.6cm} \LARGE Covariation among reproductive traits in flowering plants shape their interactions with pollinators

### 2022-10-05 Manuscript/Lanuza_et_al_2022.Rmd
Title page and MS update

---
title: \singlespacing \vspace{-1.6cm} \LARGE Covariation among reproductive traits in flowering plants shapes interactions with floral visitors

@@

Globally, plants display enormous variation in life history strategies and trait combinations. However, evidence suggests that evolutionary and physiological constraints limit the number of plant ecological strategies. Although there have been recent advances in understanding correlations among plant traits, reproductive traits are rarely considered, despite their key role in shaping plant life-history strategies and interactions with floral visitors. Here, using a global dataset of 16 reproductive traits for 1,506 species, we investigate the reproductive spectrum of flowering plants to identify how it shapes interactions with floral visitors. We show that over 50% of all trait variation is explained by the first two reproductive axes, which represent the negative correlation between flower number and flower size, and the negative correlation between autonomous selfing and floral display size. In addition, these reproductive axes were associated with different numbers of interacting partners and visitation rates of the distinct floral visitor guilds. However, the relatively small amount of variance explained by the different reproductive axes highlight the need to incorporate other factors along with reproductive traits to fully explain large scale patterns of plant-pollinator interactions. Our study identifies the major reproductive trait correlations in flowering plants and their role in shaping networks of plant-pollinator interactions at a global scale.

@@

With the recent availability of large trait databases [e.g., TRY @kattge2011;  COMPADRE @salguero2015], there has been increased research on plant ecological strategies, which has produced identification of global patterns in plant form and function [@salguero2016; @diaz2016; @bruelheide2018; @carmona2021]. However, studies that look at multiple traits tend to overlook plant reproductive traits [@evojtko2020; @ruger2018; @roddy2021], and focus on trait correlations in other plant organs such as morphological and physiological leaf traits [@wright2004; @shipley2006; @donovan2011; @osnas2013]. Despite the lack of a holistic understanding that depicts reproductive trait covariation patterns, there are widely recognized associations between plant reproductive trait pairs such as the negative correlation between flower size and flower number [@sargent2007; @kettle2011; @vojtko2022]; the positive association between flower size and outcrossing rate [@goodwillie2010]; or the positive correlation between outcrossing rate and lifespan [@barrett2003; @munoz2016; @moeller2017]). Although identification of these trait correlations has allowed progress toward a conceptual framework that integrates different floral trait relationships [@roddy2021], we still lack empirical understanding of how these associations among floral and other reproductive traits (e.g., reproductive biology) shape plant life history strategies.

Trait-based approaches [@fenster2004; @rosas2014] and trait-matching analyses [@stang2009; @bartomeus2016] are powerful for exploring the underlying mechanisms driving plant-pollinator interactions. For example, plant traits can define species' network roles [e.g., specialists vs generalists @tur2013] and plant species that occupy reproductive trait space extremes are likely to be more specialised and display tighter trait-matching with pollinators [@junker2013; @coux2016]. Indeed, the degree of morphological matching between plants and floral visitors often determines whether or not interactions occur, and consequently influences  the broader network structure [@stang2009; @ibanez2012]. However, it is still unclear how specific reproductive traits (e.g., mating or compatibility system) influence plant-pollinator associations [@sargent2008; @tur2013; @devaux2014]. Indeed, both morphology and mating system can determine a species' functional role within pollination networks, and the combination of several traits increases power for predicting network structure [@eklof2013]. Because the species position within the multidimensional trait space can result in different interaction patterns or adaptations to pollinators [see @dehling2016 for a purely morphological approach], the use of the reproductive trait space to explore plant-pollinator interactions could help to identify species functional roles and associations of traits with specific pollinators (i.e., pollination syndromes).

Here, we explore the spectrum of reproductive traits in flowering plants, within a global set of plant-pollinator networks, to identify reproductive trait covariation and how this influences the network of interactions between plants and floral visitors. First, we identify the major axes of reproductive trait variation and correlations among traits for plant species. Then, we investigate the associations between plant species’ positions in reproductive trait space and the strength of their interactions with different floral visitor guilds within networks. Finally, we determine how the main axes of reproductive trait variation influence plant species' functional roles within pollination networks, using a range of complementary interaction network metrics.

@@

All species names, genera, families and orders were retrieved and standardised from the taxonomy data sources NCBI (<https://www.ncbi.nlm.nih.gov/taxonomy>) for plants and ITIS (<https://www.itis.gov/>) for pollinators, using the R package *taxize* [@chamberlain2020]. We filled the 'not found' searches manually using <http://www.theplantlist.org/> and <http://www.mobot.org/> for plants and <http://www.catalogueoflife.org/> for floral visitors.

@@

To avoid excluding species with trait missing values, we imputed trait missing values with the help of the function *missForest* [@stekhoven2012] which allows imputation of datasets with continuous and categorical variables. We accounted for the phylogenetic distance among species on the imputation process by including the eigenvectors from principal component analysis of the phylogenetic distance (PCoA), which has been shown to improve the performance of *missForest* [@penone2014]. To extract the eigenvectors, we used the function *PVRdecomp* from the package *PVR* [@santos2018] based on a previous conceptual framework that considers phylogenetic eigenvectors [@diniz-filho2012]. We conducted two different imputations, one for the full set of species (1,506 species: 5.79% of missing values), and excluded nectar and pollen traits because of the high percentage of missing values (figure S2), and a second one for the subset of species with data for pollen per flower and microliters of nectar (755 species: 8.01% of missing values). To corroborate that our imputation of missing values did not affect our results, we evaluated the reproductive spectrum (see section below) with and without imputed values and results for the two datasets were consistent (Fig S3 and Fig S4).

@@

We used Bayesian modelling (see details below) to investigate how the different guilds of floral visitors interacted along the main axes of reproductive trait variation using qualitative and quantitative measures of plant-pollinator interactions. Specifically, interactions between plants and floral visitors in binary networks were defined as the presence or absence of floral visitation, while in weighted networks, interactions were measured as the the number of visits from each insect species to different plant species. Although floral visitors are not always pollinators and the number of visits does not consider the pollination efficiency of each species [@ballantyne2015], the number of visits does provide valuable information about the relative contributions different floral visitors make to pollination [@vazquez2005; @vazquez2012]. We divided floral visitors into six main taxonomical guilds that differ in life form, behaviour and thus likely play different ecological roles: (i) bees (Hymenoptera-Anthophila), (ii) non-bee-Hymenoptera (Hymenoptera-non-Anthophila), (iii) syrphids (Diptera-Syrphidae), (iv) non-syrphid-Diptera (Diptera-non-Syrphidae), (v) Lepidoptera, and (vi) Coleoptera. Moreover, because bees had the greatest number of occurrences in networks  (2,256 records) and the highest frequency of visits, we further broke this group down into the main bee families (Andrenidae, Apidae, Colletidae, Halictidae, and Megachilidae). In addition, we found that *Apis mellifera* was the floral visitor with the largest proportion of occurrences in networks (7.55% of the total). This is consistent with previous research showing that *A. mellifera* was the most frequent floral visitor in a similar dataset of 80 plant-pollinator networks in natural ecosystems [@hung2018]. Hence, to control for the effect of *A. mellifera* on observed visitation patterns of bees, we conducted an analogous analysis with presence-absence of the interaction and number of visits excluding *A. mellifera*. We found that *A. mellifera*, was partly driving some of the observed trends on PC1 (figure S5). However, we did not detect major differences on PC2 and PC3.

We used Bayesian generalised linear mixed models in the R package *brms* [@burkner2017] to model the presence-absence of observed interactions and number of visits as a function of the main axes of plant trait variation and their interactions with floral visitor guilds (e.g., number of visits \~ PC1 x FGs + PC2 x FGs + PC3 x FGs). Because we were interested in possible differences in visitation patterns among floral visitor guilds to plants with different reproductive strategies, we included interactions between the main axes of trait variation (PC1, PC2 and PC3) and the different floral visitor guilds in the model. We added a nested random effect of networks within the study system to capture the variation in networks among studies and within networks. Moreover, we included the phylogenetic covariance matrix as a random factor due to the possible shared evolutionary histories of species and therefore lack of independence among them. We specified for the presence or absence of an interaction and number of visits with Bernoulli and a zero inflated negative binomial distributions, respectively. The models were run with 3,000 iterations with previous 1,000 warm up iterations and with non or very weakly informative priors from the brm function so the priors had negligible influence on the results [@burkner2017]. We set delta ($\Delta$) to 0.99 to avoid divergent transitions and visualized the posterior predictive checks with the function *pp_check* using the *bayesplot* package [@gabry2019].

@@

We also used Bayesian modelling to investigate how the main axes of trait variation determined different plant species' functional roles within pollination networks. For this, we selected complementary species-level network metrics, commonly applied in bipartite network studies [@dormann2008], with clear ecological interpretations and are relevant for defining species' functional roles in networks. Specifically, the species-level metrics we included were: (i) sum of visits per plant species; (ii) normalised degree, calculated as the number of links per plant species divided by the total possible number of partners; and (iii) specialisation (d') [@bluthgen2006], which measures the deviation of an expected random choice of the available interaction partners and ranges between 0 (maximum generalisation) and 1 (maximum specialisation). Normalised degree and specialisation were calculated with the *specieslevel* function in the R package *bipartite* [@dormann2008]. We modelled each plant species metric (sum of visits, normalised degree and plant specialisation) as a function of the three main axes of trait variation (plant species level metric \~ PC1 + PC2 + PC3). For the different response variables in each model (i.e., each plant species level metric), we used different distribution families (zero inflated negative binomial for the sum of visits, weibull for normalised degree, and zero-one inflated beta for specialisation). Finally, we used the same random factors, model settings and conducted the same posterior predictive checks for each model as detailed above in the 'visitation patterns' section.

@@

The main axes of trait variation (PC1, PC2 and PC3) partly explained the presence-absence of interaction partners (conditional $R^{2}$ = 0.26; marginal $R^{2}$ = 0.20) but little of the total number of visits (conditional $R^{2}$ = 0.31; marginal $R^{2}$ = 0.06). However, we found relevant differences across the different floral visitor guilds for both the presence-absence of interactions and number of visits (figure 3). On the flower number - flower size axis, we found  that plants with many small flowers tended to have interaction partners that were Coleoptera, non-bee-Hymenoptera and Diptera (figure 3a), whereas plants with few large flowers mostly had interaction partners that were bees and Lepidoptera. We found similar trends for visitation rates on the flower number - flower size axis,  but bees and syrphids visited plant species with many small flowers more frequently (figure 3d). On the autonomous selfing - floral display axis, we found that all plant species with lower autonomous selfing had many interaction partners and recieved many visits from all floral visitor guilds (figure 3b and figure 3e). Finally, on the style length - autonomous selfing axis, plant species with short styles and low autonomous selfing had more interaction partners for all guilds, except bees which interacted more frequently with plant species with long styles and high autonomous selfing (figure 3c). Plants with long styles and high autonomous selfing interacted more frequently with Lepidoptera and non-bee-Hymenoptera (figure 3f).

@@

```{=tex}
\vspace{5mm}
\blandscape
```
```{r, echo=FALSE, message=FALSE, cache=FALSE, results='hide', warning=FALSE, fig.height=14, fig.width=27,fig.cap= "**Figure 3.** Fitted posterior estimates of the presence-absence of interaction (a, b and c) and number of visits (d, e and f) of the different floral visitor guilds in relation to the main axes of plant reproductive trait variation (PC1, PC2 and PC3). PC1 represents the flower number - flower size axis, PC2 represents the autonomous selfing - floral display axis and PC3 represents the style length - autonomous selfing axis. For visualization purposes due to large differences between number of visits from bees and the rest of the pollinator guilds, the number of visits was log-transformed (Y-axis of lower panel)."}

@@

Variance within plant species-level network metrics was poorly explained by the three main axes of trait variation (figure S9; number of visits \~ PCs, conditional $R^{2}$ = 0.11, marginal $R^{2}$ = 0.02; normalised degree \~ PCs, conditional $R^{2}$ = 0.24, marginal $R^{2}$ = 0.02; and, specialisation \~ PCs, conditional $R^{2}$ = 0.37, marginal$R^{2}$ = 0.03). Overall, the most notable trends were found on the flower number - flower size axis (PC1) and the style length - autonomous selfing axis (PC3) for number of visits and specialisation. On PC1, the number of visits per flower was higher for plant species with more flowers but was lower for plant species with larger flowers (figure S9a). However, specialisation (d’) was higher for plant species with large flowers but lower for plant species with small flowers (figure S9g). On PC3, the number of visits was lower for plants with shorter styles and lower autonomous selfing but higher for plant species with longer styles and higher autonomous selfing (figure S9c). Again, specialisation showed the opposite trend to the number of visits (figure S9i).

@@

We found strong covariation among traits within the flowering plant reproductive spectrum, which shapes interactions with floral visitors at a global scale. Despite the enormous variability in plant reproductive structures, the first two and three axes captured over 50% and 70% of trait variation, respectively, and these axes encapsulated correlations among flower number and flower size, autonomous selfing and floral display, and style length and autonomous selfing. Although explained trait variation on the first two axes was less than previous studies of morphological and physiological traits [@diaz2016; @carmona2021], our results align with the two largest and most recent studies of plant life strategies that included reproductive, form and size related traits [@salguero2016; @vojtko2022]. The different plant reproductive axes that we identified were associated with different numbers of interacting partners and visitation rates of the distinct floral visitor guilds. However, the unexplained variability found highlights the need to account for other local factors, such as species relative abundances [@bartomeus2016], along with reproductive traits, to better explain plant-pollinator associations.

Our study provides strong evidence for several widely discussed trait associations with the help of a great diversity of phylogenetic lineages (170 families and over 1,200 species). Specifically, by using a multitrait approach, we are able to link the reproductive spectrum of plant trait variation with the previously described “fast-slow continuum” in plant [@salguero2016] and animal [@healy2019] life-history strategies. This fast-slow continuum was illustrated by the associations between reproductive traits and plant height, life form and lifespan traits. For instance, although the negative correlation between flower number and flower size has received substantial attention [@worley2000; @sargent2007; @kettle2011], it has only been investigated across a few hundred Angiosperm species and generally lacking a multitrait perspective. We found support for this trait association and show that other reproductive traits (e.g., style length or ovule number) covary along this flower number - flower size axis. Our study also complements previous work linking plant breeding systems and life forms [@petit2006; @friedman2020] by showing that most species with unisexual flowers also have woody life forms and many flowers. Moreover, we identify the previously documented positive association between outcrossing rate and floral display size [@goodwillie2010], where species with low autonomous selfing levels also allocate large resource quantities to attract floral visitors.
Although we did not include floral rewards in our main analysis due to insufficient data, we did find that plant species with large floral displays and, in particular, large flowers, tended to have higher pollen and nectar quantity. Thus, we demonstrate that globally, reproductive trait covariation in flowering plants likely imposes constraints on the broader spectrum of plant trait variation.

Overall, plant reproductive traits explained interaction patterns among and within floral visitor guilds suggesting that plant reproductive strategies partly mediate these associations across spatial large scales. For example, we found that all floral visitor guilds interacted more frequently with species with large floral displays and low ability to self pollinate, which aligns with previous findings [@hegland2005; @lazaro2013; @kaiser2014]. As predicted by evolutionary life-history theory [@stebbins1970], larger investment in floral display and rewards to attract pollinators occurs when pollinator dependence is high. Indeed, the possible range of reproductive strategies in flowering plants is constrained by the cost of reproductive organs, trait associations and selective pressures exerted by different floral visitor guilds [@galen1999]. For instance, smaller flowers are visited less overall, but they are inexpensive to produce (e.g., have a low ovule number) and low visitation rate at the individual flower level can be compensated by overall visitation at the plant level by producing many flowers resulting in a larger quantity of small fruits [@kettle2011]. In addition, we found that a substantial number of species with high autonomous selfing had large styles and were associated with a high number of interactions with bees. These species seem to rely on animal pollinators as indicated by their high investment in sizable styles that can help prevent pollen competition [@ramesha2011; @ashman2013; @lanuza2021] while having reproductive assurance in their absence. Similarly, zygomorphic flowers are specialised floral structures that promote precise pollen transfer [sargent2004; @gong2009] but they were associated with high levels of autogamy. Our findings are consistent with the idea that reproductive assurance is not rare within specialised pollination systems [@fenster2007] and highlight the need to explore the timing (e.g., early versus delayed) and selfing mechanisms [see @goodwillie2018] in relation to other reproductive traits to better understand reproductive trait correlations and plant-pollinator interactions.

Because floral traits can be associated with their animal pollinators to optimise pollen transfer, the correlation between distinct suites of reproductive traits and floral visitor guilds could indicate the existence of pollination syndromes [@fenster2004; @faegri2013; @rosas2014; @dellinger2020]. For instance, bees had a higher number of interacting partners on plant species with large flowers but interacted more frequently with species with small flowers while syrphid flies showed the opposite trend. This could be reflective of differences in the ecology and behaviour between these pollinator guilds [@rader2020; @foldesi2021] that ultimately exert different selective pressures, leading divergent plant reproductive strategies. In any case, it is worth noting that local factors (e.g., species relative abundances or phenologies) likely contribute to the variability in the association between the plant reproductive spectrum and floral visitors [@vazquez2007; @encinas2012; @bartomeus2016].

Our study provides a comprehensive assessment of  reproductive trait covariation patterns in flowering plants, using a global dataset of traits for plant species within plant-pollinator networks. We show the importance of floral visitors in mediating reproductive trait correlations and identify the need to consider broader plant reproductive strategies, rather than just flower morphology, to understand patterns of plant-pollinator interactions. Further, we provide empirical evidence of reproductive trait associations that complement trait covariation studies for other plant organs, highlighting the need to include reproductive traits in the global spectrum of plant form and function. Future work incorporating underrepresented areas of the world and additional relevant reproductive traits (e.g., flower life span and colour) will be valuable for enhancing our understanding of the plant reproductive spectrum and associated plant-pollinator interactions.

>>>>

---
title: \singlespacing \vspace{-1.6cm} \LARGE Covariation among reproductive traits in flowering plants shape their interactions with pollinators

@@

Globally, plants display enormous variation in life history strategies and trait combinations. However, evidence suggests that evolutionary and physiological constraints limit the number of plant ecological strategies. Although there have been recent advances in understanding correlations among plant traits, reproductive traits are rarely considered, despite their key role in shaping plant life-history strategies and interactions with pollinators. Here, using a global dataset of 16 reproductive traits for 1,506 species, we investigate the reproductive spectrum of flowering plants to identify how it shapes interactions with pollinators. We show that over 50% of all trait variation is explained by the first two reproductive axes, which represent the negative correlation between flower number and flower size, and the negative correlation between autonomous selfing and floral display size. In addition, these reproductive axes were associated with different numbers of interacting partners and visitation rates of the distinct pollinator guilds. However, reproductive axes explain a relatively small amount of variance in pollinator interactions highlighting the need to incorporate other factors along with reproductive traits to fully explain large scale patterns of plant-pollinator interactions. Our study identifies the major reproductive trait correlations in flowering plants and their role in shaping networks of plant-pollinator interactions at a global scale.

@@

With the recent availability of large trait databases [e.g., TRY @kattge2011;  COMPADRE @salguero2015], there has been increased research on plant ecological strategies, which has produced identification of global patterns in plant form and function [@salguero2016; @diaz2016; @bruelheide2018; @carmona2021]. However, studies that look at multiple traits tend to overlook plant reproductive traits [@evojtko2020; @ruger2018; @roddy2021], and focus on trait correlations in other plant organs such as morphological and physiological leaf traits [@wright2004; @shipley2006; @donovan2011; @osnas2013]. Despite the lack of a holistic understanding that depicts reproductive trait covariation patterns, there are widely recognized associations between plant reproductive trait pairs such as the negative correlation between flower size and flower number [@sargent2007; @kettle2011; @vojtko2022]; the positive association between flower size and outcrossing rate [@goodwillie2010]; or the positive correlation between outcrossing rate and lifespan [@barrett2003; @munoz2016; @moeller2017]. Although identification of these trait correlations has allowed progress toward a conceptual framework that integrates different floral trait relationships [@roddy2021], we still lack empirical understanding of how these associations among floral and other reproductive traits (e.g., reproductive biology) shape plant life history strategies.

Trait-based approaches [@fenster2004; @rosas2014] and trait-matching analyses [@stang2009; @bartomeus2016] are powerful for exploring the underlying mechanisms driving plant-pollinator interactions. For example, plant traits can define species' network roles [e.g., specialists vs generalists @tur2013] and plant species that occupy reproductive trait space extremes are likely to be more specialised and display tighter trait-matching with pollinators [@junker2013; @coux2016]. Indeed, the degree of morphological matching between plants and pollinators often determines whether or not interactions occur, and consequently influences  the broader network structure [@stang2009; @ibanez2012]. However, it is still unclear how specific reproductive traits (e.g., mating or compatibility system) influence plant-pollinator associations [@sargent2008; @tur2013; @devaux2014]. For example, species with high selfing rates may evolve to invest less in floral rewards resulting in a lower number of pollinator visits [@devaux2014]. Indeed, both morphology and mating system can determine a species' functional role within pollination networks, and the combination of several traits increases power for predicting network structure [@eklof2013]. Because the species position within the multidimensional trait space can result in different interaction patterns or adaptations to pollinators [see @dehling2016 for a purely morphological approach], the use of the reproductive trait space to explore plant-pollinator interactions could help to identify species functional roles and associations of traits with specific pollinators (i.e., pollination syndromes).

Here, we explore the spectrum of reproductive traits in flowering plants, within a global set of plant-pollinator networks, to identify reproductive trait covariation and how this influences the network of interactions between plants and pollinators. First, we identify the major axes of reproductive trait variation and correlations among traits for plant species. Then, we investigate the associations between plant species’ positions in reproductive trait space and the strength of their interactions with different pollinator guilds within networks. Finally, we determine how the main axes of reproductive trait variation influence plant species' functional roles within pollination networks, using a range of complementary interaction network metrics.

@@

All species names, genera, families and orders were retrieved and standardised from the taxonomy data sources NCBI (<https://www.ncbi.nlm.nih.gov/taxonomy>) for plants and ITIS (<https://www.itis.gov/>) for pollinators, using the R package *taxize* [@chamberlain2020]. We filled the 'not found' searches manually using <http://www.theplantlist.org/> and <http://www.mobot.org/> for plants and <http://www.catalogueoflife.org/> for pollinators.

@@

To avoid excluding species with trait missing values, we imputed trait missing values with the help of the function *missForest* [@stekhoven2012] which allows imputation of datasets with continuous and categorical variables. We accounted for the phylogenetic distance among species on the imputation process by including the eigenvectors from principal component analysis of the phylogenetic distance (PCoA), which has been shown to improve the performance of *missForest* [@penone2014]. To extract the eigenvectors, we used the function *PVRdecomp* from the package *PVR* [@santos2018] based on a previous conceptual framework that considers phylogenetic eigenvectors [@diniz-filho2012]. We conducted two different imputations, one for the full set of species (1,506 species: 5.79% of missing values), and excluding nectar and pollen traits because of the high percentage of missing values (figure S2), and a second one for the subset of species with data for pollen per flower and microliters of nectar (755 species: 8.01% of missing values). To corroborate that our imputation of missing values did not affect our results, we evaluated the reproductive spectrum (see section below) with and without imputed values and results for the two datasets were consistent (Fig S3 and Fig S4).

@@

We used Bayesian modelling (see details below) to investigate how the different guilds of pollinators interacted along the main axes of reproductive trait variation using qualitative and quantitative measures of plant-pollinator interactions. Specifically, interactions between plants and pollinators in binary networks were defined as the presence or absence of floral visitation, while in weighted networks, interactions were measured as the the number of visits from each insect species to different plant species. Although floral visitors are not always pollinators and the number of visits does not consider the pollination efficiency of each species [@ballantyne2015], the number of visits does provide valuable information about the relative contributions different floral visitors make to pollination [@vazquez2005; @vazquez2012]. Therefore, despite the fact that we do not evaluate pollinator effectiveness we refer to the different floral visitors as pollinators.

We divided pollinators into six main taxonomical guilds that differ in life form, behaviour and thus likely play different ecological roles: (i) bees (Hymenoptera-Anthophila), (ii) non-bee-Hymenoptera (Hymenoptera-non-Anthophila), (iii) syrphids (Diptera-Syrphidae), (iv) non-syrphid-Diptera (Diptera-non-Syrphidae), (v) Lepidoptera, and (vi) Coleoptera. Moreover, because bees had the greatest number of occurrences in networks  (2,256 records) and the highest frequency of visits, we further broke this group down into the main bee families (Andrenidae, Apidae, Colletidae, Halictidae, and Megachilidae). In addition, we found that *Apis mellifera* was the pollinator with the largest proportion of occurrences in networks (7.55% of the total). This is consistent with previous research showing that *A. mellifera* was the most frequent pollinator in a similar dataset of 80 plant-pollinator networks in natural ecosystems [@hung2018]. Hence, to control for the effect of *A. mellifera* on observed visitation patterns of bees, we conducted an analogous analysis with presence-absence of the interaction and number of visits excluding *A. mellifera*. We found that *A. mellifera*, was partly driving some of the observed trends on PC1 (figure S5). However, we did not detect major differences on PC2 and PC3.

We used Bayesian generalised linear mixed models in the R package *brms* [@burkner2017] to model the presence-absence of observed interactions and number of visits as a function of the main axes of plant trait variation and their interactions with pollinator guilds (e.g., number of visits \~ PC1 x FGs + PC2 x FGs + PC3 x FGs). Because we were interested in possible differences in visitation patterns among pollinator guilds to plants with different reproductive strategies, we included interactions between the main axes of trait variation (PC1, PC2 and PC3) and the different pollinator guilds in the model. We added a nested random effect of networks within the study system to capture the variation in networks among studies and within networks. Moreover, we included the phylogenetic covariance matrix as a random factor due to the possible shared evolutionary histories of species and therefore lack of independence among them. We specified for the presence or absence of an interaction and number of visits with Bernoulli and a zero inflated negative binomial distributions, respectively. The models were run with 3,000 iterations with previous 1,000 warm up iterations and with non or very weakly informative priors from the brm function so the priors had negligible influence on the results [@burkner2017]. We set delta ($\Delta$) to 0.99 to avoid divergent transitions and visualised the posterior predictive checks with the function *pp_check* using the *bayesplot* package [@gabry2019].

@@

We also used Bayesian modelling to investigate how the main axes of trait variation determined different plant species' functional roles within pollination networks. For this, we selected complementary species-level network metrics, commonly applied in bipartite network studies [@dormann2008], with clear ecological interpretations and are relevant for defining species' functional roles in networks. Specifically, the species-level metrics we included were: (i) sum of visits per plant species; (ii) normalised degree, calculated as the number of links per plant species divided by the total possible number of partners; and (iii) selectivity (d') [@bluthgen2006], which measures the deviation of an expected random choice of the available interaction partners and ranges between 0 (maximum generalisation) and 1 (maximum specialisation). Normalised degree and selectivity were calculated with the *specieslevel* function in the R package *bipartite* [@dormann2008]. We modelled each plant species metric (sum of visits, normalised degree and plant selectivity) as a function of the three main axes of trait variation (plant species level metric \~ PC1 + PC2 + PC3). For the different response variables in each model (i.e., each plant species level metric), we used different distribution families (zero inflated negative binomial for the sum of visits, weibull for normalised degree, and zero-one inflated beta for selectivity). Finally, we used the same random factors, model settings and conducted the same posterior predictive checks for each model as detailed above in the 'visitation patterns' section.

@@

The main axes of trait variation (PC1, PC2 and PC3) partly explained the presence-absence of interaction partners (conditional $R^{2}$ = 0.26; marginal $R^{2}$ = 0.20) but little of the total number of visits (conditional $R^{2}$ = 0.31; marginal $R^{2}$ = 0.06). However, we found relevant differences across the different pollinator guilds for both the presence-absence of interactions and number of visits (figure 3). On the flower number - flower size axis, we found  that plants with many small flowers tended to have interaction partners that were Coleoptera, non-bee-Hymenoptera and Diptera (figure 3a), whereas plants with few large flowers mostly had interaction partners that were bees and Lepidoptera. We found similar trends for visitation rates on the flower number - flower size axis,  but bees and syrphids visited plant species with many small flowers more frequently (figure 3d). On the autonomous selfing - floral display axis, we found that all plant species with lower autonomous selfing had many interaction partners and recieved many visits from all pollinator guilds (figure 3b and figure 3e). Finally, on the style length - autonomous selfing axis, plant species with short styles and low autonomous selfing had more interaction partners for all guilds, except bees which interacted more frequently with plant species with long styles and high autonomous selfing (figure 3c). Plants with long styles and high autonomous selfing interacted more frequently with Lepidoptera and non-bee-Hymenoptera (figure 3f).

@@

```{=tex}
\vspace{5mm}
\blandscape
```
```{r, echo=FALSE, message=FALSE, cache=FALSE, results='hide', warning=FALSE, fig.height=14, fig.width=27,fig.cap= "**Figure 3.** Fitted posterior estimates of the presence-absence of interaction (a, b and c) and number of visits (d, e and f) of the different pollinator guilds in relation to the main axes of plant reproductive trait variation (PC1, PC2 and PC3). PC1 represents the flower number - flower size axis, PC2 represents the autonomous selfing - floral display axis and PC3 represents the style length - autonomous selfing axis. For visualization purposes due to large differences between number of visits from bees and the rest of the pollinator guilds, the number of visits was log-transformed (Y-axis of lower panel)."}

@@

Variance within plant species-level network metrics was poorly explained by the three main axes of trait variation (figure S9; number of visits \~ PCs, conditional $R^{2}$ = 0.11, marginal $R^{2}$ = 0.02; normalised degree \~ PCs, conditional $R^{2}$ = 0.24, marginal $R^{2}$ = 0.02; and, selectivity \~ PCs, conditional $R^{2}$ = 0.37, marginal$R^{2}$ = 0.03). Overall, the most notable trends were found on the flower number - flower size axis (PC1) and the style length - autonomous selfing axis (PC3) for number of visits and selectivity. On PC1, the number of visits per flower was higher for plant species with more flowers but was lower for plant species with larger flowers (figure S9a). However, selectivity (d’) was higher for plant species with large flowers but lower for plant species with small flowers (figure S9g). On PC3, the number of visits was lower for plants with shorter styles and lower autonomous selfing but higher for plant species with longer styles and higher autonomous selfing (figure S9c). Again, selectivity showed the opposite trend to the number of visits (figure S9i).

@@

We found strong covariation among traits within the flowering plant reproductive spectrum, which shapes interactions with pollinators at a global scale. Despite the enormous variability in plant reproductive structures, the first two and three axes captured over 50% and 70% of trait variation, respectively, and these axes encapsulated correlations among flower number and flower size, autonomous selfing and floral display, and style length and autonomous selfing. Although explained trait variation on the first two axes was less than previous studies of morphological and physiological traits [@diaz2016; @carmona2021], our results align with the two largest and most recent studies of plant life strategies that included reproductive, form and size related traits [@salguero2016; @vojtko2022]. The different plant reproductive axes that we identified were associated with different numbers of interacting partners and visitation rates of the distinct pollinator guilds. However, the unexplained variability found highlights the need to account for other local factors, such as species relative abundances [@bartomeus2016], along with reproductive traits, to better explain plant-pollinator associations.

Our study provides strong evidence for several widely discussed trait associations with the help of a great diversity of phylogenetic lineages (170 families and over 1,200 species). Specifically, by using a multitrait approach, we are able to link the reproductive spectrum of plant trait variation with the previously described “fast-slow continuum” in plant [@salguero2016] and animal [@healy2019] life-history strategies. This fast-slow continuum was illustrated by the associations between reproductive traits and plant height, life form and lifespan traits. For instance, although the negative correlation between flower number and flower size has received substantial attention [@worley2000; @sargent2007; @kettle2011], it has only been investigated across a few hundred Angiosperm species and generally lacking a multitrait perspective. We found support for this trait association and show that other reproductive traits (e.g., style length or ovule number) covary along this flower number - flower size axis. Our study also complements previous work linking plant breeding systems and life forms [@petit2006; @friedman2020] by showing that most species with unisexual flowers also have woody life forms and many flowers. Moreover, we identify the previously documented positive association between outcrossing rate and floral display size [@goodwillie2010], where species with low autonomous selfing levels also allocate large resource quantities to attract pollinators.

Although we did not include floral rewards in our main analysis due to insufficient data, we did find that plant species with large floral displays and, in particular, large flowers, tended to have higher pollen and nectar quantity. Thus, we demonstrate that globally, reproductive trait covariation in flowering plants likely imposes constraints on the broader spectrum of plant trait variation.

Overall, plant reproductive traits explained interaction patterns among and within pollinator guilds suggesting that plant reproductive strategies partly mediate these associations across spatial large scales. For example, we found that all pollinator guilds interacted more frequently with species with large floral displays and low ability to self pollinate, which aligns with previous findings [@hegland2005; @lazaro2013; @kaiser2014]. As predicted by evolutionary life-history theory [@stebbins1970], larger investment in floral display and rewards to attract pollinators occurs when pollinator dependence is high. Indeed, the possible range of reproductive strategies in flowering plants is constrained by the cost of reproductive organs, trait associations and selective pressures exerted by different pollinator guilds [@galen1999]. For instance, smaller flowers are visited less overall, but they are inexpensive to produce (e.g., have a low ovule number) and low visitation rate at the individual flower level can be compensated by overall visitation at the plant level by producing many flowers resulting in a larger quantity of small fruits [@kettle2011]. In addition, we found that a substantial number of species with high autonomous selfing had large styles and were associated with a high number of interactions with bees. These species seem to rely on animal pollinators as indicated by their high investment in sizable styles that can help prevent pollen competition [@ramesha2011; @ashman2013; @lanuza2021] while having reproductive assurance in their absence. For instance, the annual herb _Scabiosa atropurpurea_ has flowers grouped in clusters with relatively long styles and a single ovule per flower that is likely to be fertilised with a single legitimate floral visit but avoids reproductive failure with autonomous self pollination. Similarly, zygomorphic flowers are specialised floral structures that promote precise pollen transfer [@sargent2004; @gong2009] but they were associated with high levels of autogamy. Our findings are consistent with the idea that reproductive assurance is not rare within specialised pollination systems [@fenster2007] and highlight the need to explore the timing (e.g., early versus delayed) and selfing mechanisms [see @goodwillie2018] in relation to other reproductive traits to better understand reproductive trait correlations and plant-pollinator interactions.

Because floral traits can be associated with their animal pollinators to optimise pollen transfer, the correlation between distinct suites of reproductive traits and pollinator guilds could indicate the existence of pollination syndromes [@fenster2004; @faegri2013; @rosas2014; @dellinger2020]. For instance, bees had a higher number of interacting partners on plant species with large flowers but interacted more frequently with species with small flowers while syrphid flies showed the opposite trend. This could be reflective of differences in the ecology and behaviour between these pollinator guilds [@rader2020; @foldesi2021] that ultimately exert different selective pressures, leading divergent plant reproductive strategies. In any case, it is worth noting that local factors (e.g., species relative abundances or phenologies) likely contribute to the variability in the association between the plant reproductive spectrum and pollinators [@vazquez2007; @encinas2012; @bartomeus2016]. Taking both aspects together, we suggest that while plant reproductive traits do shape their interactions with pollinators, pollinator syndromes at a macroecological scale are not distinct compartmentalised items, but form part of a multifaceted context dependent continuum.

Our study provides a comprehensive assessment of  reproductive trait covariation patterns in flowering plants, using a global dataset of traits for plant species within plant-pollinator networks. We show the importance of pollinators in mediating reproductive trait correlations and identify the need to consider broader plant reproductive strategies, rather than just flower morphology, to understand patterns of plant-pollinator interactions. Further, we provide empirical evidence of reproductive trait associations that complement trait covariation studies for other plant organs, highlighting the need to include reproductive traits in the global spectrum of plant form and function. Future work incorporating underrepresented areas of the world and additional relevant reproductive traits (e.g., flower life span and colour) will be valuable for enhancing our understanding of the plant reproductive spectrum and associated plant-pollinator interactions.

\newpage
